Tree size, exposure, and hydraulic traits interactively shape drought response in a temperate broadleaf forest

### 2019-06-28 manuscript/03-methods.Rmd
Added images
Added code to make tables

![Structure of ForestGEO plot](tables_figures/ForestGEO_plot.jpg)

@@

### Biophysical traits
Canopy classes were observed in the field during the growing season of 2018 following the crown position protocol from [@jennings_assessing_1999], whereby positions were ranked as dominant, codominant, intermediate, or suppressed. For ease of core analysis, these labels were condensed to canopy (including dominant and codominant) and subcanopy (intermediate and suppressed).

@@

Need to say the methodology for all traits used
1. PLA
2. LMA
3. Chlorophyll
4. Wood density
5. TLP
6. P50 = psi_0.5_kl50
7. P80 = psi_0.5_kl80

```{r, echo=FALSE}
library(knitr)
library(kableExtra)

traits_table <- data.frame(
  "Trait" = c("Ring Porosity", "Percent Leaf Area", "Leaf Mass Area", "Chlorophyll", "Wood density", "TLP", "P50", "P80"),
  "Unit" = c("", "%", "g/m2", "m2/g", "g/cm3", "MPa", "MPa", "MPa"))

@@

>>>>

![Structure of ForestGEO plot](tables_figures/ForestGEO_plot.jpg){width=500px}

@@

![Time series of combined tree cores by species](tables_figures/Time_series_for_each_species.jpg){width=500px}

### Biophysical traits
Canopy classes were observed in the field during the growing season of 2018 following the crown position protocol from [@jennings_assessing_1999], whereby positions were ranked as dominant, codominant, intermediate, or suppressed. In determining pointer years, these labels were condensed to canopy (including dominant and codominant) and subcanopy (intermediate and suppressed). The full set of canopy classes were used for all other analyses.

```{r, Table-cores-position, eval=TRUE, echo=FALSE, message=FALSE}
library(RCurl)
library(dplyr)
library(tidyr)
library(data.table)
library(knitr)
library(kableExtra)

species <- read.csv(text=getURL("https://raw.githubusercontent.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/master/data/core_list_for_neil.csv?token=AJNRBEIAVYFAN2CJB3Y3UAS5D5BNG"), stringsAsFactors = FALSE)

table <- species[, c(1:3,6,17)]

positions <- read.csv(text=getURL("https://raw.githubusercontent.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/master/tree_dimensions/tree_crowns/cored_dendroband_crown_position_data/dendro_cored_full.csv"))

table$position_all <- positions$crown.position[match(table$tag, positions$tag)]
table$position_all <- gsub("D", "dominant", table$position_all)
table$position_all <- gsub("C", "co-dominant", table$position_all)
table$position_all <- gsub("I", "intermediate", table$position_all)
table$position_all <- gsub("S", "suppressed", table$position_all)

table1 <- table %>%
  group_by(sp) %>%
  summarize(n_cores=n())

table2 <- table %>%
  group_by(sp, position_all) %>%
  summarize(count=n())

table2 <- spread(table2, key=position_all, value=count)
setnames(table2, old="<NA>", new="prior dead")

table2$n_cores <- table1$n_cores[match(table2$sp, table1$sp)]
table2 <- table2[, c(1,7,3,2,4:6)]
table2 <- table2[!table2$sp %in% c("frni", "pist"), ]

kable(table2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

@@

Need to say the methodology for all traits used (**Nobby making short list**)
1. PLA
2. LMA
3. Chlorophyll
4. Wood density
5. TLP
6. P50 = psi_0.5_kl50
7. P80 = psi_0.5_kl80

```{r, traits-overview, eval=TRUE, echo=FALSE, message=FALSE}
library(knitr)
library(kableExtra)

trees_all <- read.csv("tables_figures/trees_all.csv", stringsAsFactors = FALSE)

trees_all <- trees_all[!duplicated(trees_all$sp), ]

traits_table <- data.frame(
  "Trait" = c("Ring Porosity", "Percent Leaf Area", "Leaf Mass Area", "Chlorophyll", "Wood density", "TLP", "P50", "P80"),
  "Unit" = c("ring, semi-ring, diffuse", "%", "g/m2", "m2/g", "g/cm3", "MPa", "MPa", "MPa"))
traits_table$abb <- c("rp", "PLA", "LMA", "Chl", "WD", "TLP", "p50", "p80")

traits_table$mean <- NA
traits_table$min <- NA
traits_table$max <- NA

for (i in seq(along=traits_table$abb[2:8])){
  trait <- traits_table$abb[2:8][i]

  sub <- trees_all[grepl(trait, colnames(trees_all))]
  traits_table$mean[2:8][i] <- ifelse(grepl(trait, colnames(sub)), 
                              mean(sub[, 1], na.rm=TRUE), 
                              traits_table$mean)

  traits_table$min[2:8][i] <- ifelse(grepl(trait, colnames(sub)), 
                              min(sub[, 1], na.rm=TRUE), 
                              traits_table$min)

  traits_table$max[2:8][i] <- ifelse(grepl(trait, colnames(sub)), 
                              max(sub[, 1], na.rm=TRUE), 
                              traits_table$max)
}

traits_table[, c("mean", "min", "max")] <- 
  sapply(traits_table[, c("mean", "min", "max")], function(x)
  round(x, 2)
)
traits_table$abb <- NULL

@@

## Model runs
Traits' importance in predicting drought tolerance was calculated from mixed-effects models and the lowest AICc [@R-lme4, @R-AICcmodavg]. Specifically, we first determined the most important biophysical traits before including them in the model with the leaf hydraulic traits [**need to say why?**]

### 2019-06-28 manuscript/04-results.Rmd
Added images
Tried to make correct output formats
Added code to make tables

# Results

## Biophysical Models
Crown position 
When including only biophysical traits, trees' resistance value (on a per-species basis) is explained best by crown position and height, with codominant trees being the most resistant to drought. This follows on work done by [@bennett_larger_2015] [and others?] which show that larger trees suffer more during drought, and confirms that this susceptibility can be seen in tree ring analyses.

@@

>>>>

```{r include=FALSE, cache=FALSE}
rm(list = ls(all = TRUE))
```
# Results

## Model runs
When including only biophysical traits, trees' resistance value (on a per-species basis) is explained best by crown position and height, with codominant trees being the most resistant to drought. This follows on work done by [@bennett_larger_2015] [and others?] which show that larger trees suffer more during drought, and confirms that this susceptibility can be seen in tree ring analyses. Adding in crown position with the leaf hydraulic traits yields a slightly worse predictive model for drought tolerance, with height remaining as the only significant biophysical variable.

``` {r, biophysical-model, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE}
library(lme4)
library(AICcmodavg) #aictab function
library(car)
library(data.table)
library(knitr)
library(kableExtra)

trees_all_bio <- read.csv("tables_figures/trees_all_bio.csv", stringsAsFactors = FALSE)

response <- "resist.value"
effects <- c("position_all", "elev.m", "distance.ln.m", "height.ln.m", "year", "(1|sp)")

#create all combinations of random / fixed effects
effects_comb <- 
  unlist( sapply( seq_len(length(effects)), 
                  function(i) {
                    apply( combn(effects,i), 2, function(x) paste(x, collapse = "+"))
                  }))

# pair response with effect and sub out combinations that don't include random effects
#in general, if two variables are >70% correlated, you can toss one of them without significantly affecting the results
var_comb <- expand.grid(response, effects_comb) 
var_comb <- var_comb[grepl("1", var_comb$Var2), ] #only keep in fixed/random combos
var_comb <- var_comb[grepl("year", var_comb$Var2), ] #keep year in for drought sake

# formulas for all combinations. $Var1 is the response, and $Var2 is the effect
# for good stats, you should have no more total parameters than 1/10th the number of observations in your dataset
formula_vec <- sprintf("%s ~ %s", var_comb$Var1, var_comb$Var2)

# create list of model outputs
lmm_all <- lapply(formula_vec, function(x){
  fit1 <- lmer(x, data = trees_all_bio, REML=FALSE, 
               control = lmerControl(optimizer ="Nelder_Mead"))
  return(fit1)
})
names(lmm_all) <- formula_vec

var_aic <- aictab(lmm_all, second.ord=TRUE, sort=TRUE) #rank based on AICc

best <- lmm_all[[8]]
cof_bio <- data.frame("value" = coef(summary(best))[ , "Estimate"])
cof_bio <- setDT(cof_bio, keep.rownames = TRUE)[]
setnames(cof_bio, old="rn", new="coefficients")

```
``` {r, full-model, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE}
trees_all_full <- read.csv("tables_figures/trees_all_full.csv", stringsAsFactors = FALSE)

response <- "resist.value"
effects <- c("rp", "PLA_dry_percent", "LMA_g_per_m2", "WD_g_per_cm3", "mean_TLP_Mpa", "height.ln.m", "position_all", "year", "(1|sp/tree)")

#create all combinations of random / fixed effects
effects_comb <- 
  unlist( sapply( seq_len(length(effects)), 
                  function(i) {
                    apply( combn(effects,i), 2, function(x) paste(x, collapse = "+"))
                  }))

# pair response with effect and sub out combinations that don't include random effects
#in general, if two variables are >70% correlated, you can toss one of them without significantly affecting the results
var_comb <- expand.grid(response, effects_comb) 
var_comb <- var_comb[grepl("1", var_comb$Var2), ] #only keep in fixed/random combos
var_comb <- var_comb[grepl("year", var_comb$Var2), ] #keep year in for drought sake

# formulas for all combinations. $Var1 is the response, and $Var2 is the effect
# for good stats, you should have no more total parameters than 1/10th the number of observations in your dataset
formula_vec <- sprintf("%s ~ %s", var_comb$Var1, var_comb$Var2)

# create list of model outputs
lmm_all <- lapply(formula_vec, function(x){
  fit1 <- lmer(x, data = trees_all_full, REML=FALSE, 
               control = lmerControl(optimizer ="Nelder_Mead"))
  return(fit1)
})
names(lmm_all) <- formula_vec

var_aic <- aictab(lmm_all, second.ord=TRUE, sort=TRUE) #rank based on AICc

best <- lmm_all[[121]]
cof_full <- data.frame("value" = coef(summary(best))[ , "Estimate"])
cof_full <- setDT(cof_full, keep.rownames = TRUE)[]
setnames(cof_full, old="rn", new="coefficients")
```
```{r, make-tables, eval=TRUE, echo=FALSE, message=FALSE, results = 'asis'}
library(knitr)
library(kableExtra)

cof_bio %>%
  kable("html", align = 'clc', caption = 'Biophysical Model') %>%
    kable_styling(full_width = F, position = "float_left")

cof_full %>%
  kable("html", align = 'clc', caption = 'Full Model') %>%
    kable_styling(full_width = F, position = "right")
```

@@

### Vertical Profiles

![Climate variables with height](tables_figures/NEON_vertical_profiles.png){width=500px}
